---
title: "**Homework 8**"
author: 
- "**Name:** Sofia Vaquera"
- "**UT EID:** sv27974"
- "**Github:** https://github.com/svaquera/HW-8"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loads the tidyverse and mosaic libraries
library(tidyverse)
library(mosaic)

```

# Problem 1: Regression Warm Up

```{r, echo = FALSE}

# Imports creatinine dataset
creatinine = read.csv("creatinine.csv")

```


## A) Predict Creatinine Clearance Rate

```{r, echo = FALSE}

# Fits regression model to data and returns coefficients
model_creatinine = lm(creatclear ~ age, data = creatinine)
coef(model_creatinine)

# Fitted regression model is about 147.813 - 0.620x


# Calculates predicted creatinine clearance rate for a 55 year old
147.813 - (0.620 * 55)


```



The creatinine clearance rate one would expect for a 55-year-old is about 113.713 mL/minute. This was determined by fitting a regression model to the data, using the `age` variable as the feature or "predictor" variable for `creatclear`, the target or "response" variable. We used the coefficients obtained from this process, which were 147.813 for the intercept and -0.620 for the weight on `age`, and created the fitted regression model y hat = 147.813 - 0.620x. After plugging in the value 55 (the individual's age) for x, the equation becomes y hat = 147.813 - (0.620 * 55), which is about 113.713 mL/minute.


## B) Change in Creatinine Clearance with Age
```{r, echo = FALSE}

# Returns coefficients for fitted regression model
coef(model_creatinine)

```
Creatinine clearance rate decreases by about 0.620 mL/minute per year. This was determined by examining the coefficients of the fitted regression model calculated above in Part A, which show that for each year that age increases, creatinine clearance rate decreases by about 0.620 mL/minute.



## C) Relative Creatinine Clearance Rates
```{r, echo = FALSE}

# Equation: 147.813 - 0.620x

# Calculates 40-year-old's predicted creatinine clearance rate
predicted_40 = 147.813 - (0.620 * 40)
predicted_40

# Calculates residual (actual - predicted)
residual_40 = 135 - predicted_40
residual_40


# Calculates 60-year-old's predicted creatinine clearance rate
predicted_60 = 147.813 - (0.620 * 60)
predicted_60

# Calculates residual (actual - predicted)
residual_60 = 112 - predicted_60
residual_60

```

The 40-year-old has a healthier (higher) creatinine clearance rate for their age. This was determined by examining the residuals (actual creatinine clearance rate - predicted creatinine clearance rate) for both the 40-year-old and the 60-year-old. It is given that the 40-year-old has an actual creatinine clearance rate of 135 mL/minute, and we can find their predicted creatinine clearance rate by plugging 40 (the individual's age) into the fitted regression model calculated in Part A, meaning that the 40-year-old's predicted creatinine clearance rate is about 147.813 - (0.620 * 40) = 123.013. By subtracting the 40-year-old's predicted creatinine clearance rate from their actual creatinine clearance rate, we find the 40-year-old's residual to be 135 - 123.013 = 11.987 mL/minute. It is given that the 60-year-old has an actual creatinine clearance rate of 112 mL/minute, and we can find their predicted creatinine clearance rate by plugging 60 (the individual's age) into the fitted regression model calculated in Part A, meaning that the 60-year-old's predicted creatinine clearance rate is about 147.813 - (0.620 * 60) = 110.613. By subtracting the 60-year-old's predicted creatinine clearance rate from their actual creatinine clearance rate, we find the 60-year-old's residual to be 112 - 110.613 = 1.387 mL/minute. Given that the 40-year-old has a creatinine clearance rate that's about 11.987 mL/minute above what is expected for their age while the 60-year-old has a creatinine clearance rate that's only about 1.387 mL/minute above what is expected for their age, the 40-year old has the healthier (higher) creatinine clearance rate for their age.


## Problem 2: Modeling Disease Growth

```{r, echo = FALSE}

# Imports covid dataset
covid = read.csv("covid.csv")

# Creates dataset using only values from Italy
italy = covid |>
  filter(country == "Italy")

# Creates dataset using only values from Spain
spain = covid |>
  filter(country == "Spain")

```

## A) Italy's Growth Rate and Doubling Time
```{r, message = FALSE, echo = FALSE}

# Fits regression model to data and calculates rate and doubling time
lm_italy = lm(log(deaths) ~ days_since_first_death, data = italy)
italy_rate = coef(lm_italy)[["days_since_first_death"]]
italy_doubling_time = 70 / (italy_rate * 100)

# Prints rate and doubling time
italy_rate
italy_doubling_time


# Sets seed for reproducibility
set.seed(315)

# Creates bootstrap for growth rate and creates 95% confidence interval 
boot_italy = do(10000) * lm(log(deaths) ~ days_since_first_death, data = resample(italy))
confint(boot_italy, level = 0.95)

# Creates bootstrap for doubling time and creates 95% confidence interval
boot_italy_doubling_time = numeric(10000)
for (i in 1:10000){
  sample_lm = lm(log(deaths) ~ days_since_first_death, data = resample(italy))
  sample_rate = (coef(sample_lm)["days_since_first_death"]) * 100
  boot_italy_doubling_time[i] = 70 / sample_rate
}
confint(boot_italy_doubling_time, level = 0.95)


```
The estimated growth rate for Italy is 0.183, or about 18.3%, while the estimated doubling time is about 3.8 days. The 95% confidence interval for the growth rate is between 0.159 and 0.209, while the 95% confidence interval for the doubling time is between 3.4 and 4.4.


## B) Spain's Growth Rate and Doubling Time
```{r, message = FALSE, echo = FALSE}

# Fits regression model to data and calculates rate and doubling time
lm_spain = lm(log(deaths) ~ days_since_first_death, data = spain)
spain_rate = coef(lm_spain)[["days_since_first_death"]]
spain_doubling_time = 70 / (spain_rate * 100)

# Prints rate and doubling time
spain_rate
spain_doubling_time


# Sets seed for reproducibility
set.seed(315)

# Creates bootstrap for growth rate and creates 95% confidence interval 
boot_spain = do(10000) * lm(log(deaths) ~ days_since_first_death, data = resample(spain))
confint(boot_spain, level = 0.95)

# Creates bootstrap for doubling time and creates 95% confidence interval
boot_spain_doubling_time = numeric(10000)
for (i in 1:10000){
  sample_lm = lm(log(deaths) ~ days_since_first_death, data = resample(spain))
  sample_rate = (coef(sample_lm)["days_since_first_death"]) * 100
  boot_spain_doubling_time[i] = 70 / sample_rate
}
confint(boot_spain_doubling_time, level = 0.95)

```

The estimated growth rate for Spain is 0.276, or about 27.6%, while the estimated doubling time is about 2.5 days. The 95% confidence interval for the growth rate is between 0.235 and 0.317, while the 95% confidence interval for the doubling time is between 2.2 and 3.0.


## C) Line Graph of Reported Daily Deaths Over Time By Country
```{r, echo = FALSE}

# Creates a line graph showing reported daily deaths over time in each country
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) + geom_line() + 
  labs(title = "Reported Daily COVID Deaths Over Time in Italy & Spain",
       x = "Days Since First Death", 
       y = "Reported Daily Deaths")

```

# Problem 3: Price Elasticity of Demand

```{r, echo = FALSE}

# Imports milk dataset
milk = read.csv("milk.csv")

# Fits regression model to data and returns coefficients
lm_milk = lm(log(sales) ~ log(price), data = milk)
coef(lm_milk)


# Sets seed for reproducibility
set.seed(315)

# Creates bootstrap for price elasticity and creates 95% confidence interval 
boot_milk = do(10000) * lm(log(sales) ~ log(price), data = resample(milk))
confint(boot_milk, level = 0.95)


```

The estimated price elasticity of demand for milk is about -1.619. The 95% confidence interval for price elasticity is between -1.772 and -1.457. In order to estimate the elasticity, it was identified that there is a relative proportional change in y versus x (sales versus price), which allowed us to fit a linear model on a logarithmic scale for the x and y variables. In other words, we were able to determine the elasticity by fitting a power law model using linear regression to log(y) vs log(x) (which in this case would be log(sales) versus log(price)). After examining the coefficients of the fitted model, we took the fitted slope to represent elasticity; hence, we found the value of -1.619 to be the elasticity, representing that for every 1% the price of milk increases, the amount that people want to purchase decreases by about 1.619%. We repeated the process of creating a linear model and examining its coefficients to determine elasticity 10000 times to create bootstrapped samples, allowing us to calculate that the 95% bootstrapped confidence interval is between -1.772 and -1.457.
 